---
title: "Power simulations for sensitivity analyses (Umemoto et al. preregistered report)"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, results='show', warning=FALSE, message=FALSE, echo=TRUE, include=TRUE}
library(data.table); library(hausekeep); library(tidyverse)
```

## Simulations

We simulated 3888 different parameter combinations for our linear mixed model `y ~ b0 + b1x + b2x + error`. For each set of parameter combination, we simulated 1000 datasets and fitted the models using the `lmer()` function from the `lme4` package and obtained the significance values using the `lmerTest` package. Therefore, we ran 3888 * 1000 = 3888000 models in total. For each test, we determine whether the probability value was equal to or smaller than 0.02 (> 0.02 coded as 0, <= 0.02 coded as 1). These simulations were run using the package `paramtest`. 

Simulated random-intercept model: `lmer(y ~ x + y + (1 | subject)`

Since the raw simulated data are nearly 1GB in size, we did not upload them here. We instead provide only the aggregated dataset whereby each row reflects the aggregated values/estimates from 1000 simulations using the same set of parameters. For simulation script, see `simPower.R` script.

Data key

* `N.test`: simulated sample size (60, 70, 80, 90)
* `b1.test`: simulated b1 coefficients (0.05, 0.08, 0.12)
* `b2.test`: simulated b1 coefficients (0.05, 0.08, 0.12)
* `varInt.test`: simulated variance in intercept b0 across participants (0.01, 0.05, 0.10)
* `varSlope_b1.test`: simulated variance in b1 across participants (0.01, 0.05, 0.10)
* `varSlope_b2.test`: simulated variance in b2 across participants (0.01, 0.05, 0.10)
* `varResid.test`: simulated residual variance (0.01, 0.10, 0.20, 0.30)
* `power_b1`: observed power (1000 simulations) (alpha = 0.02)
* `power_b2`: observed power (1000 simulations) (alpha = 0.02)
* `r_x1`: mean observed effect size correlation r (1000 simulations)
* `r_x2`: mean observed effect size correlation r (1000 simulations)

### Read data

```{r read data}
dt1 <- fread("power_simulations_avg.csv")
print(dt1)
```

### Compute minimum observed power

```{r}
avg <- dt1[, .(pow_b1 = mean(power_b1), pow_b2 = mean(power_b2), 
               min_r_x1 = min(r_x1), min_r_x2 = min(r_x2)), 
        keyby = .(N.test, b1.test)]
# convert correlation r to Cohen's f and print result
avg[, f := es(r = min_r_x1, msg = F)$f, by = .(N.test, b1.test)] %>% print()
```

### Plot

```{r power sensitivity plot}
ggplot(avg, aes(f, pow_b1, col = as.factor(N.test))) +
  geom_point() +
  # geom_line() +
  geom_hline(yintercept = 0.90, linetype = 'dashed') +
  labs(x = "Mininum simulated effect size (Cohen's f)", y = "Statistical power (alpha = 0.02)", col = "Sample size")
ggsave("simulations.jpg", dpi = 200, width = 7, height = 5)
```


